---
title: "Covariate analysis of CMC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (ACC and DLPFC with CQN normlaisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_MSSM.Rmd",
                                 parentId = "syn7501880",
                                 entityName = 'ACC and DLPFC - STAR - FEATURE COUNTS - ENSEMBL v75 - CQN - MSSM')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(reshape2)
library(limma)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(cqn)

library(synapseClient)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

synapseLogin()

library(doParallel)
library(foreach)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = "syn7501880";
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of STAR aligned feature counts with Ensembl v75 with CQN normalisation (ACC and DLPFC)';

thisFileName <- 'CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_MSSM.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='CMC')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download reprocessed counts (ACC)
COUNT_ACC_ID = 'syn8414424'
ALL_USED_IDs = COUNT_ACC_ID
COUNT_ACC = downloadFile(COUNT_ACC_ID) 

# Download reprocessed counts (DLPFC)
COUNT_DLPFC_ID = 'syn8413221';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = COUNT_DLPFC_ID
COUNT_DLPFC = downloadFile(COUNT_DLPFC_ID)

# Join ACC and DLPFC together
COUNT = join_all(list(COUNT_ACC, COUNT_DLPFC), type = 'full') %>%
  as.data.frame()

GENE.LEN = dplyr::select(COUNT_ACC, Geneid, Length) %>%
  dplyr::rename(ensembl_gene_id = Geneid) %>%
  unique()
rownames(GENE.LEN) = GENE.LEN$ensembl_gene_id

rownames(COUNT) = COUNT$Geneid
ind = setdiff(colnames(COUNT), c('Geneid', 'Chr', 'Start', 'End', 'Strand', 'Length'))
COUNT = data.matrix(COUNT[,ind])

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY = read.table(synGet(ANCESTRY_ID)@filePath, fill = T, header = T, sep = '\t') %>%
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'GenotypingSampleID'))

# Get RNA qc metrics from synapse
METADATA_QC_ACC_ID = 'syn2929053'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_QC_ACC_ID
METADATA_QC_ACC = read.table(synGet(METADATA_QC_ACC_ID)@filePath, fill = T, header = T, sep = ',') %>%
  lapply(as.character)

# Get RNA qc metrics from synapse
METADATA_QC_DLPFC_ID = 'syn2279446'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_QC_DLPFC_ID
METADATA_QC_DLPFC = read.table(synGet(METADATA_QC_DLPFC_ID)@filePath, fill = T, header = T, sep = ',') %>%
  lapply(as.character)

METADATA_QC = rbindlist(list(METADATA_QC_ACC, METADATA_QC_DLPFC), use.names = T, fill = T) %>%
  dplyr::filter(is.na(Exclude.)) %>%
  dplyr::select(Sample.RNA.ID, Ribozero.Batch, Library.Batch, Flowcell.Batch, Mapped.Reads, 
                Intragenic.Rate, Exonic.Rate, Intronic.Rate, Intergenic.Rate, Genes.Detected, 
                Transcripts.Detected, Expression.Profiling.Efficiency, rRNA.Rate, Total.Reads, Percent.Aligned) %>%
  dplyr::rename(SampleID = Sample.RNA.ID, RibozeroBatch = Ribozero.Batch, LibraryBatch = Library.Batch, 
                FlowcellBatch = Flowcell.Batch, MappedReads = Mapped.Reads, IntragenicRate = Intragenic.Rate, 
                ExonicRate = Exonic.Rate, IntronicRate = Intronic.Rate, IntergenicRate = Intergenic.Rate, 
                GenesDetected = Genes.Detected, TranscriptsDetected = Transcripts.Detected, 
                ExpProfEfficiency = Expression.Profiling.Efficiency, rRNARate = rRNA.Rate, 
                TotalReads = Total.Reads, PercentAligned = Percent.Aligned)

# Get merged metadata
METADATA_ID = 'syn4484273'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_ID
METADATA = read.csv(synGet(METADATA_ID)@filePath, fill = T, header = T, check.names = F)

## Metadata with specific information for RNASeq processing
metadataVarOfInterest = c("Individual ID", "Institution", "Gender", "Age of Death", "PMI (in hours)", "Dx", 
                      
                          "ACC_RNA_isolation: Exclude?", "ACC_RNA_isolation: Sample RNA ID",
                          "ACC_RNA_isolation: RIN", "ACC_RNA_report: Exclude?", 
                          
                          "DLPFC_RNA_isolation: Exclude?", "DLPFC_RNA_isolation: Sample RNA ID",
                          "DLPFC_RNA_isolation: RIN", "DLPFC_RNA_report: Exclude?",
                          
                          "DNA_genotyping_report: Genotyping Sample_ID")

# Filter metadata 
ind = colnames(METADATA) %in% metadataVarOfInterest
METADATA = METADATA[,ind]
colnames(METADATA) = gsub('[^[:alnum:]]','_',colnames(METADATA))

METADATA_ACC = METADATA %>%
  dplyr::select(Individual_ID, Institution, Gender, Age_of_Death, PMI__in_hours_, Dx, 
                DNA_genotyping_report__Genotyping_Sample_ID, ACC_RNA_isolation__Exclude_, 
                ACC_RNA_isolation__Sample_RNA_ID, ACC_RNA_isolation__RIN, ACC_RNA_report__Exclude_) %>%
  dplyr::rename(PMI = PMI__in_hours_, 
                IsolationExclude = ACC_RNA_isolation__Exclude_, 
                SampleID = ACC_RNA_isolation__Sample_RNA_ID, 
                RIN = ACC_RNA_isolation__RIN, 
                ReportExclude = ACC_RNA_report__Exclude_,
                GenotypingSampleID = DNA_genotyping_report__Genotyping_Sample_ID) %>%
  dplyr::filter(SampleID %in% colnames(COUNT)) %>%
  dplyr::mutate(Tissue = 'ACC')

METADATA_DLPFC = METADATA %>%
  dplyr::select(Individual_ID, Institution, Gender, Age_of_Death, PMI__in_hours_, Dx, 
                DNA_genotyping_report__Genotyping_Sample_ID, DLPFC_RNA_isolation__Exclude_, 
                DLPFC_RNA_isolation__Sample_RNA_ID, DLPFC_RNA_isolation__RIN, DLPFC_RNA_report__Exclude_) %>%
  dplyr::rename(PMI = PMI__in_hours_, 
                IsolationExclude = DLPFC_RNA_isolation__Exclude_, 
                SampleID = DLPFC_RNA_isolation__Sample_RNA_ID, 
                RIN = DLPFC_RNA_isolation__RIN, 
                ReportExclude = DLPFC_RNA_report__Exclude_,
                GenotypingSampleID = DNA_genotyping_report__Genotyping_Sample_ID) %>%
  dplyr::filter(SampleID %in% colnames(COUNT)) %>%
  dplyr::mutate(Tissue = 'DLPFC')

METADATA = rbindlist(list(METADATA_ACC, METADATA_DLPFC), use.names = T, fill = T) %>%
  droplevels() %>%
  list(.,METADATA_QC) %>%
  join_all %>%
  dplyr::mutate(LibraryBatch = as.factor(LibraryBatch)) %>%
  left_join(ANCESTRY) %>%
  dplyr::filter(Institution == 'MSSM') %>%
  unique

# Replace AFF and BP samples as other
levels(METADATA$Dx) = c("Other", "Other", "Control", "SCZ")
```

### Data preprocessing
```{r preprocess.data}
# Remove samples marked as exclude and samples with no Gender, Ethnicity, PMI, and RIN
writeLines('Following counts are missing any metadata')
writeLines(paste(setdiff(colnames(COUNT), levels(METADATA$SampleID)), collapse = ', '))
METADATA <- METADATA %>% filter(SampleID %in% colnames(COUNT)) 

writeLines('Following samples are marked exclude')
writeLines(paste(METADATA$SampleID[METADATA$IsolationExclude == 1 | METADATA$ReportExclude == 1], collapse = ', '))
METADATA <- METADATA  %>% filter(IsolationExclude == 0, ReportExclude == 0) 

writeLines('Following samples are missing PMI information')
writeLines(paste(METADATA$SampleID[is.na(METADATA$PMI)], collapse = ', '))
METADATA <- METADATA  %>% filter(!is.na(PMI)) 

writeLines('Following samples are missing gender information')
writeLines(paste(METADATA$SampleID[is.na(METADATA$Gender)], collapse = ', '))
METADATA <- METADATA  %>% filter(!is.na(Gender)) 

writeLines('Following samples are missing RIN information')
writeLines(paste(METADATA$SampleID[is.na(METADATA$RIN)], collapse = ', '))
METADATA <- METADATA  %>% filter(!is.na(RIN)) 

writeLines('Following samples are missing EV information')
writeLines(paste(METADATA$SampleID[is.na(METADATA$EV.1)], collapse = ', '))
METADATA <- METADATA  %>% filter(!is.na(METADATA$EV.1)) 
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(COUNT))
removedIDs = setdiff(colnames(COUNT), METADATA$SampleID)

COUNT = COUNT[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx, Tissue) %>% 
  summarise(count = n()) %>% 
  spread(Tissue, count) %>%
  kable()

METADATA = METADATA %>%
  dplyr::mutate(Dx.Tissue = paste(Dx, Tissue, sep = '.'))
```
Following sample are removed `r paste(removedIDs, collapse = ',')`

```{r gene.param}
## Get GC content from biomart
backgroundGenes = data.frame(ensembl_gene_id = rownames(COUNT))

# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "dec2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "percentage_gc_content", "gene_biotype"),
                       filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
                       mart = mart)

GENE.GC.CONT = Ensemble2HGNC %>%
  dplyr::select(ensembl_gene_id, percentage_gc_content) %>%
  unique
rownames(GENE.GC.CONT) = GENE.GC.CONT$ensembl_gene_id
```

### Covariates clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c("Individual_ID", "Gender", "LibraryBatch", "Dx.Tissue", "RibozeroBatch", "FlowcellBatch")
ContCovariates <- c("Age_of_Death", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
                    "GenesDetected", "TranscriptsDetected", "ExpProfEfficiency", "rRNARate", "TotalReads", 
                    "PercentAligned", "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
rownames(COVARIATES) <- METADATA$SampleID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
tmp = COVARIATES.CORRELATION$ESTIMATE
tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0
h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate')
draw(h, heatmap_legend_side = 'left')
```

### Explore metatdata
```{r data.explore, fig.width = 12, fig.height = 12}
my.theme <- theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1), plot.title=element_text(hjust=0.5))

# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = RIN)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme

# Age of Death
p[[2]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = Age_of_Death)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + my.theme

# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = PMI)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI (in hours)') + my.theme

# Intronic Rate
p[[4]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntronicRate)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Intronic Rate') + my.theme

# IntergenicRate
p[[5]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntergenicRate)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Intergenic Rate') + my.theme

# Transcripts Detected
p[[6]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = TranscriptsDetected)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Transcripts Detected') + my.theme

# Mapped Reads
p[[7]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = MappedReads)) + geom_boxplot()
p[[7]] = p[[7]] + ggtitle('Mapped Reads') + my.theme

# PercentAligned
p[[8]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = PercentAligned)) + geom_boxplot()
p[[8]] = p[[8]] + ggtitle('Percent Aligned') + my.theme

# rRNARate
p[[9]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = rRNARate)) + geom_boxplot()
p[[9]] = p[[9]] + ggtitle('rRNARate') + my.theme

multiplot(plotlist = p, cols = 3)

# Institution
# vcd::mosaic(~ Institution + Dx.Tissue, data = COVARIATES)

# Gender
# vcd::mosaic(~ Gender + Dx.Tissue, data = COVARIATES)
```
### Filter genes
Remove genes that have less than 1 cpm counts in at least 50% of samples per Dx.Tissue. Also remove genes with missing gene length and percentage GC content
```{r cpmnormalisation}
genesToAnalyze = dlply(METADATA, .(Dx.Tissue), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID %>% droplevels()],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC.CONT$ensembl_gene_id[!is.na(GENE.GC.CONT$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$Length)])

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
pct.pc = PROCESSED_COUNTS$filteredExprMatrix$genes %>%
  dplyr::rename(ensembl_gene_id = genes) %>%
  left_join(Ensemble2HGNC) %>%
  group_by(gene_biotype) %>%
  summarise(fraction  = n()) %>%
  filter(fraction > 100) %>%
  dplyr::mutate(fraction = fraction/length(PROCESSED_COUNTS$filteredExprMatrix$genes[,1]))
```
`r dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]` genes are used for the analysis. 

Following is the distribution of biotypes
`r kable(pct.pc)`

### Outlier analysis
Detect outlier samples based on their expression (logCPM) pattern
```{r detect.outliers, fig.height=8, fig.width=8, results='asis', cache = FALSE}
# Based on PC
indToRemove = c('MSSM_RNA_ACC_BP_24', 'MSSM_RNA_ACC_375', 'MSSM_RNA_ACC_170', 'MSSM_RNA_ACC_315', 'MSSM_RNA_ACC_250',
                'MSSM_RNA_ACC_21', 'MSSM_RNA_ACC_356', 'MSSM_RNA_ACC_318', 'MSSM_RNA_ACC_222', 'MSSM_RNA_ACC_220',
                'MSSM_RNA_ACC_379', 'MSSM_RNA_PFC_133', 'MSSM_RNA_PFC_139', 'MSSM_RNA_PFC_163', 'MSSM_RNA_PFC_260',
                'MSSM_RNA_PFC_299', 'MSSM_RNA_PFC_361', 'MSSM_RNA_PFC_89', 'MSSM_RNA_PFC_319', 'MSSM_RNA_PFC_348',
                'MSSM_RNA_PFC_354', 'MSSM_RNA_PFC_60', 'PENN_RNA_PFC_71', 'PENN_RNA_PFC_56',  "MSSM_RNA_ACC_310",
                "MSSM_RNA_ACC_51",  "MSSM_RNA_ACC_254", "MSSM_RNA_ACC_28", "MSSM_RNA_ACC_293", "MSSM_RNA_ACC_233",
                "MSSM_RNA_ACC_167", "MSSM_RNA_ACC_353", "MSSM_RNA_ACC_355", "MSSM_RNA_ACC_211", "PENN_RNA_ACC_122",
                "PENN_RNA_ACC_049", "PENN_RNA_ACC_052", "PENN_RNA_ACC_068", "PENN_RNA_ACC_070", "PITT_RNA_ACC_10020",
                "PITT_RNA_ACC_1240", "PITT_RNA_ACC_1284", "PITT_RNA_ACC_700", "PITT_RNA_ACC_857", "PITT_RNA_ACC_BP_1589",
                "MSSM_RNA_PFC_230", "MSSM_RNA_PFC_304", "MSSM_RNA_PFC_309", "MSSM_RNA_PFC_329", "MSSM_RNA_PFC_82",
                "PENN_RNA_PFC_5", "PENN_RNA_PFC_58", "PENN_RNA_PFC_99", "PITT_RNA_PFC_1284")

# # Outliers from Menahcem's DLPFC analysis syn3493755
# outliers = c('MSSM_RNA_BP_PFC_1', 'MSSM_RNA_PFC_21', 'MSSM_RNA_BP_PFC_3', 'MSSM_RNA_PFC_103', 
#              'MSSM_RNA_PFC_105', 'MSSM_RNA_PFC_115', 'MSSM_RNA_PFC_169',  'MSSM_RNA_PFC_17',
#              'MSSM_RNA_PFC_18', 'MSSM_RNA_PFC_19', 'MSSM_RNA_PFC_213', 'MSSM_RNA_PFC_214', 
#              'MSSM_RNA_PFC_215', 'MSSM_RNA_PFC_216', 'MSSM_RNA_PFC_217', 'MSSM_RNA_PFC_218',
#              'MSSM_RNA_PFC_219', 'MSSM_RNA_PFC_222', 'MSSM_RNA_PFC_24', 'MSSM_RNA_PFC_252', 
#              'MSSM_RNA_PFC_259', 'MSSM_RNA_PFC_263', 'MSSM_RNA_PFC_267', 'MSSM_RNA_PFC_275',
#              'MSSM_RNA_PFC_307', 'MSSM_RNA_PFC_322', 'MSSM_RNA_PFC_325', 'MSSM_RNA_PFC_27', 
#              'MSSM_RNA_PFC_332', 'MSSM_RNA_PFC_333', 'MSSM_RNA_PFC_334', 'MSSM_RNA_PFC_335',
#              'MSSM_RNA_PFC_340', 'MSSM_RNA_PFC_35', 'MSSM_RNA_PFC_362', 'MSSM_RNA_PFC_372', 
#              'MSSM_RNA_PFC_373', 'MSSM_RNA_PFC_385', 'MSSM_RNA_PFC_393', 'MSSM_RNA_PFC_395',
#              'MSSM_RNA_PFC_49', 'MSSM_RNA_PFC_54', 'MSSM_RNA_PFC_70', 'MSSM_RNA_PFC_76', 
#              'MSSM_RNA_PFC_77', 'MSSM_RNA_PFC_78', 'MSSM_RNA_PFC_88', 'MSSM_RNA_PFC_93',
#              'FAKE_DLPFC_CMC_MSSM_426', 'PENN_RNA_PFC_29', 'FAKE_DLPFC_CMC_PENN_055', 
#              'PENN_RNA_PFC_75', 'PENN_RNA_PFC_110', 'FAKE_DLPFC_CMC_PENN_106', 'FAKE_DLPFC_CMC_PENN_107',
#              'FAKE_DLPFC_CMC_PENN_108', 'MSSM_RNA_PFC_14', 'MSSM_RNA_BP_PFC_21', 'MSSM_RNA_PFC_293', 
#              'MSSM_RNA_PFC_349', 'MSSM_RNA_PFC_353', 'MSSM_RNA_PFC_363', 'MSSM_RNA_PFC_369',
#              'PENN_RNA_PFC_82', 'PENN_RNA_PFC_89', 'PITT_RNA_PFC_1159', 'PITT_RNA_PFC_1296', 'PITT_RNA_PFC_537', 
#              'PITT_RNA_PFC_567', 'PITT_RNA_PFC_871', 'PITT_RNA_PFC_878', 'MSSM_RNA_PFC_133', 'MSSM_RNA_PFC_139',
#              'MSSM_RNA_PFC_210',  'MSSM_RNA_PFC_260', 'MSSM_RNA_PFC_278', 'MSSM_RNA_PFC_299', 'MSSM_RNA_PFC_348',
#              'MSSM_RNA_PFC_361', 'MSSM_RNA_PFC_60', 'PENN_RNA_PFC_56', 'PITT_RNA_BP_PFC_10003', 'PITT_RNA_BP_PFC_1047',
#              'PITT_RNA_BP_PFC_1086', 'PITT_RNA_BP_PFC_1092', 'PITT_RNA_BP_PFC_1247', 'PITT_RNA_BP_PFC_1282', 
#              'PITT_RNA_BP_PFC_1324', 'PITT_RNA_BP_PFC_1374', 'PITT_RNA_BP_PFC_1391', 'PITT_RNA_BP_PFC_686')

# Find principal components of expression to plot
PC <- prcomp(voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
  dplyr::mutate(label = SampleID)
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Tissue, size=Age_of_Death))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Dx+.~Tissue, scale = 'free')
p

indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]

tmp = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue'), sep = '_') %>%
  group_by(Dx, Tissue) %>%
  summarise(count = n()) %>%
  spread(Tissue, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples from `r length(unique(COVARIATES$Individual_ID))` unique individuals

Based on the expression pattern following `r length(indToRemove)` samples were tagged as outliers based on their mrna expression: `r paste(indToRemove, collapse = ', ')` 

Distribution of samples are:
`r kable(tmp)`

### Winsorise counts (Gene outliers in samples)
```{r winsorise.data}
LOG.CPM = cpm(PROCESSED_COUNTS$filteredExprMatrix$counts, log = T)

# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(LOG.CPM, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  
  ind.low = which(x == max(x[which(x <= mn-3*std.dev)], na.rm = T))
  ind.high = which(x == min(x[which(x >= mn+3*std.dev)], na.rm = T))
  
  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t

# Back calculate counts for further analysis
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Library Normalisation
Initial library normalisation usign cqn
```{r cqn}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(NEW.COUNTS, 
                          x = GENE.GC.CONT[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gc_content'],
                          lengths = GENE.LEN[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'Length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset

# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
CQN.GENE_EXPRESSION$E = apply(CQN.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  
  ind.low = which(x == max(x[which(x <= mn-3*std.dev)], na.rm = T))
  ind.high = which(x == min(x[which(x >= mn+3*std.dev)], na.rm = T))
  
  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t

# Back calculate counts for further analysis
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^CQN.GENE_EXPRESSION$E) * t(replicate(dim(CQN.GENE_EXPRESSION$E)[1], LIB.SIZE))/1e6
```

### Co-expression distribution
Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(CQN.GENE_EXPRESSION$E))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Sample clustering
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_')

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Tissue, size=Age_of_Death))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
p
```
Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Gender", "LibraryBatch", "Dx.Tissue")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```
```{r temp, include = F}
dev.off()
gc()
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE}
# Plot abberent distribution of logcpm counts
tmp1 = (CQN.GENE_EXPRESSION$E) %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(Dx+.~Tissue, scale = 'free')
p
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preAdjustedSigCovars.NULL, fig.width=20, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation
Based on different model selection methodology (as seen here syn8468116, syn9772932, and syn9774491), following variables are used as covariates: Gender, RIN, and IntronicRate

NOTE: Dx.Tissue is chosen as the primary variable of interest (i.e., covariate selection is conditioned on diagnosis and tissue), and Individual\_ID is chosen as the random effect

```{r custom.normalisation, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Dx.Tissue', 'Gender', 'RIN', 'IntronicRate')

writeLines(paste('Using following covariates in the model:',
                 paste(postAdjustCovars, collapse=', '),
                 'as fixed effects and Individual_ID is chosen as random effect'))
  
# Post adjusted design matrix
DESIGN = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)$design
DESIGN = DESIGN[,linColumnFinder(DESIGN)$indepCols]
  
# Estimate voom weights
VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, design = DESIGN)
VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E
correlation <- parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION, block = COVARIATES$Individual_ID)
  
# Re-calculate voom weights with correlation of random effects
VOOM.GENE_EXPRESSION = voom(NEW.COUNTS,   
                            design = DESIGN, plot=F,
                            block = COVARIATES$Individual_ID, 
                            correlation = correlation$cor)
VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E

# Fit linear model using new weights and new design
ADJUSTED.FIT = lmFit(VOOM.GENE_EXPRESSION)
  
# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(ADJUSTED.FIT, CQN.GENE_EXPRESSION$E)
```

### Residual calculation
Calculate weighted residuals and add back "Dx.Tissue" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("Dx.Tissue", colnames(DESIGN), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
  ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DESIGN[,varsToAddIn])
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
  dplyr::mutate(label = SampleID)
# plotdata$label[!(plotdata$SampleID %in% top.nsamples)] = ''

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Tissue, size=Age_of_Death))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
# p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
```

```{r decompse.normalise.data2.1, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx.Tissue'),drop=F])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp[,c('Dx.Tissue'),drop=F]))
```

```{r temp1, include=F}
dev.off()
gc()
```

### Differential expression analysis (SCZ - Control)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=15, fig.width=20}
# Fit contrast
contrast = makeContrasts(contrasts=c("Dx.TissueSCZ_ACC-Dx.TissueControl_ACC",
                                     "Dx.TissueSCZ_ACC-Dx.TissueOther_ACC",
                                     "Dx.TissueOther_ACC-Dx.TissueControl_ACC",
                                     
                                     "Dx.TissueSCZ_DLPFC-Dx.TissueControl_DLPFC",
                                     "Dx.TissueSCZ_DLPFC-Dx.TissueOther_DLPFC",
                                     "Dx.TissueOther_DLPFC-Dx.TissueControl_DLPFC",
                                     
                                     "Dx.TissueSCZ_DLPFC-Dx.TissueSCZ_ACC",
                                     "Dx.TissueControl_DLPFC-Dx.TissueControl_ACC",
                                     "Dx.TissueOther_DLPFC-Dx.TissueOther_ACC",
                                     
                                     "Dx.TissueSCZ_DLPFC+Dx.TissueSCZ_ACC-Dx.TissueControl_DLPFC-Dx.TissueControl_ACC",         
                                     "Dx.TissueSCZ_DLPFC+Dx.TissueControl_DLPFC-Dx.TissueSCZ_ACC-Dx.TissueControl_ACC",
                                     "Dx.TissueSCZ_DLPFC-Dx.TissueControl_DLPFC-Dx.TissueSCZ_ACC+Dx.TissueControl_ACC"),
                         levels = colnames(ADJUSTED.FIT$coefficients))
FIT.CONTR = contrasts.fit(ADJUSTED.FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = TRUE) %>%
    rownameToFirstColumn('ensembl_gene_id') 
}, FIT.CONTR) 
names(DE) = gsub('Dx.Tissue', '', colnames(contrast))
names(DE)[10:12] = c('SCZ_ALL-Control_ALL', 'ALL_DLPFC-ALL_ACC', 'ALL_ALL-ALL_ALL')

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Dx.Tissue','',Comparison),
                Direction = sign(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, into = c('to.state','ref.state'), sep = '-') %>%
  tidyr::separate(to.state, into = c('to.Dx','to.Tissue'), sep = '_') %>%
  tidyr::separate(ref.state, into = c('ref.Dx','ref.Tissue'), sep = '_') %>%
  left_join(Ensemble2HGNC) %>%
  left_join(GENE.LEN)
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::filter(adj.P.Val <= 0.05, ref.Dx != 'Other', to.Dx != 'Other') %>%
  dplyr::select(ensembl_gene_id, to.Dx, to.Tissue, ref.Dx, ref.Tissue, logFC) %>%
  group_by(to.Dx, to.Tissue, ref.Dx, ref.Tissue) %>%
  dplyr::summarise(FDR_0_05_DOWN = length(unique(ensembl_gene_id[logFC < 0])),
                   FDR_0_05_UP = length(unique(ensembl_gene_id[logFC > 0])))

tmp1 = DE %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2), ref.Dx != 'Other', to.Dx != 'Other') %>%
  dplyr::select(ensembl_gene_id, to.Dx, to.Tissue, ref.Dx, ref.Tissue, logFC) %>%
  group_by(to.Dx, to.Tissue, ref.Dx, ref.Tissue) %>%
  dplyr::summarise(FDR_0_05_FC_1.2_DOWN = length(unique(ensembl_gene_id[logFC < 0])),
                   FDR_0_05_FC_1.2_UP = length(unique(ensembl_gene_id[logFC > 0])))

kable(full_join(tmp,tmp1))

p = DE %>%
  dplyr::filter(ref.Dx != 'ALL', to.Dx != 'ALL') %>%
  ggplot(aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position = 'top')
p = p + facet_grid(ref.Dx+ref.Tissue~.+to.Dx+to.Tissue, scales = 'fixed')
p = p + ggtitle('Differential Expression between Schizophrenia and Control')
p

p = DE %>%
  dplyr::filter(ref.Dx == 'ALL', to.Dx == 'ALL') %>%
  ggplot(aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position = 'top')
p = p + facet_grid(ref.Tissue+.~to.Tissue, scales = 'fixed')
p = p + ggtitle('Differential Expression between DLPFC and ACC')
p
```

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=10, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE %>% dplyr::filter(ref.Dx == 'Control', to.Dx == 'SCZ'), 
                 aes(x = log10(Length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(Length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(ref.Tissue~.) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')

pl[[2]] = ggplot(DE %>% dplyr::filter(ref.Dx == 'Control', to.Dx == 'SCZ'), 
                 aes(x = percentage_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(ref.Tissue~.) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')

pl[[3]] = ggplot(DE %>% dplyr::filter(ref.Dx == 'Control', to.Dx == 'SCZ'), 
                 aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(ref.Tissue~.) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')

multiplot(plotlist = pl, cols = 3)
```

### Differential expression analysis (Gender specific SCZ - Control)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp1, fig.height=15, fig.width=20}
# Covariates to adjust
COVARIATES$Dx.Tissue.Gender = factor(paste(COVARIATES$Dx.Tissue, COVARIATES$Gender, sep = '_'))
postAdjustCovars = c('Dx.Tissue.Gender', 'RIN', 'IntronicRate')

writeLines(paste('Using following covariates in the model:',
                 paste(postAdjustCovars, collapse=', '),
                 'as fixed effects and Individual_ID is chosen as random effect'))
  
# Post adjusted design matrix
DESIGN1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)$design
DESIGN1 = DESIGN1[,linColumnFinder(DESIGN1)$indepCols]
  
# Estimate voom weights
VOOM.GENE_EXPRESSION1 = voom(NEW.COUNTS, design = DESIGN1)
VOOM.GENE_EXPRESSION1$E = CQN.GENE_EXPRESSION$E
correlation <- parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION1, block = COVARIATES$Individual_ID)
  
# Re-calculate voom weights with correlation of random effects
VOOM.GENE_EXPRESSION1 = voom(NEW.COUNTS,   
                             design = DESIGN1, plot=F,
                             block = COVARIATES$Individual_ID, 
                             correlation = correlation$cor)
VOOM.GENE_EXPRESSION1$E = CQN.GENE_EXPRESSION$E

# Fit linear model using new weights and new design
ADJUSTED.FIT1 = lmFit(VOOM.GENE_EXPRESSION1)

# Fit contrast
contrast = makeContrasts(contrasts=c("Dx.Tissue.GenderSCZ_ACC_Male-Dx.Tissue.GenderControl_ACC_Male",
                                     "Dx.Tissue.GenderSCZ_ACC_Female-Dx.Tissue.GenderControl_ACC_Female",
                                     
                                     "Dx.Tissue.GenderSCZ_DLPFC_Male-Dx.Tissue.GenderControl_DLPFC_Male",
                                     "Dx.Tissue.GenderSCZ_DLPFC_Female-Dx.Tissue.GenderControl_DLPFC_Female"),
                         levels = colnames(ADJUSTED.FIT1$coefficients))
FIT.CONTR1 = contrasts.fit(ADJUSTED.FIT1, contrasts=contrast)
FIT.CONTR1 = eBayes(FIT.CONTR1)

# Get differnetial expression
DE1 = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = TRUE) %>%
    rownameToFirstColumn('ensembl_gene_id') 
}, FIT.CONTR1) 
names(DE1) = gsub('Dx.Tissue.Gender', '', colnames(contrast)) 

DE1 = DE1 %>% 
  data.table::rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Dx.Tissue.Gender','',Comparison),
                Direction = sign(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, into = c('to.state','ref.state'), sep = '-') %>%
  tidyr::separate(to.state, into = c('to.Dx', 'to.Tissue', 'to.Gender'), sep = '_') %>%
  tidyr::separate(ref.state, into = c('ref.Dx', 'ref.Tissue', 'ref.Gender'), sep = '_') %>%
  left_join(Ensemble2HGNC) %>%
  left_join(GENE.LEN)
DE1$Direction[DE1$adj.P.Val > 0.05 | abs(DE1$logFC) < log2(1.2)] = 'NONE'

tmp = DE1 %>%
  dplyr::filter(adj.P.Val <= 0.05) %>%
  dplyr::select(ensembl_gene_id, to.Dx, to.Tissue, to.Gender, ref.Dx, ref.Tissue, ref.Gender, logFC) %>%
  group_by(to.Dx, to.Tissue, to.Gender, ref.Dx, ref.Tissue, ref.Gender) %>%
  dplyr::summarise(FDR_0_05_DOWN = length(unique(ensembl_gene_id[logFC < 0])),
                   FDR_0_05_UP = length(unique(ensembl_gene_id[logFC > 0])))

tmp1 = DE1 %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) %>%
  dplyr::select(ensembl_gene_id, to.Dx, to.Tissue, to.Gender, ref.Dx, ref.Tissue, ref.Gender, logFC) %>%
  group_by(to.Dx, to.Tissue, to.Gender, ref.Dx, ref.Tissue, ref.Gender) %>%
  dplyr::summarise(FDR_0_05_FC_1.2_DOWN = length(unique(ensembl_gene_id[logFC < 0])),
                   FDR_0_05_FC_1.2_UP = length(unique(ensembl_gene_id[logFC > 0])))

kable(full_join(tmp,tmp1))

p = DE1 %>%
  ggplot(aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position = 'top')
p = p + facet_grid(ref.Tissue~.+ref.Gender, scales = 'fixed')
p = p + ggtitle('Differential Expression between Schizophrenia and Control')
p

table(COVARIATES$Dx.Tissue, COVARIATES$Gender)
```

### Differential expression analysis (ACC - DLPFC, SCZ - Control, Female - Male)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp2, fig.height=16, fig.width=8}
COVARIATES = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue')) %>%
  dplyr::mutate(Dx = factor(Dx, levels = c('SCZ', 'Control', 'Other')),
                Tissue = factor(Tissue))
postAdjustCovars = c('Tissue', 'Dx', 'Gender', 'RIN', 'IntronicRate')

# Post adjusted design matrix
DESIGN2 = getDesignMatrix(droplevels(COVARIATES[,postAdjustCovars,drop=F]),Intercept = T)$design
DESIGN2 = DESIGN2[,linColumnFinder(DESIGN2)$indepCols]
  
# Estimate voom weights
VOOM.GENE_EXPRESSION2 = voom(NEW.COUNTS, design = DESIGN2)
VOOM.GENE_EXPRESSION2$E = CQN.GENE_EXPRESSION$E
correlation <- parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION2, block = COVARIATES$Individual_ID)
  
# Re-calculate voom weights with correlation of random effects
VOOM.GENE_EXPRESSION2 = voom(NEW.COUNTS,   
                             design = DESIGN2, plot=F,
                             block = COVARIATES$Individual_ID, 
                             correlation = correlation$cor)
VOOM.GENE_EXPRESSION2$E = CQN.GENE_EXPRESSION$E

# Fit linear model using new weights and new design
ADJUSTED.FIT2 = lmFit(VOOM.GENE_EXPRESSION2)

# Fit contrast
contrast = makeContrasts(contrasts=c("TissueDLPFC",
                                     "DxControl",
                                     "GenderMale"),
                         levels = colnames(ADJUSTED.FIT2$coefficients))
FIT.CONTR2 = contrasts.fit(ADJUSTED.FIT2, contrasts=contrast)
FIT.CONTR2 = eBayes(FIT.CONTR2)

# Get differnetial expression
DE2 = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = TRUE) %>%
    rownameToFirstColumn('ensembl_gene_id') 
}, FIT.CONTR2) 
names(DE2) = c('ACC-DLPFC', 'SCZ-CONTROL', 'FEMALE-MALE')

DE2 = DE2 %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Direction = sign(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  left_join(Ensemble2HGNC) 
DE2$Direction[DE2$adj.P.Val > 0.05 | abs(DE2$logFC) < log2(1.2)] = 'NONE'

tmp = DE2 %>%
  dplyr::filter(adj.P.Val <= 0.05) %>%
  dplyr::select(Comparison, ensembl_gene_id, logFC) %>%
  group_by(Comparison) %>%
  dplyr::summarise(FDR_0_05_DOWN = length(unique(ensembl_gene_id[logFC < 0])),
                   FDR_0_05_UP = length(unique(ensembl_gene_id[logFC > 0])))

tmp1 = DE2 %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) %>%
  dplyr::select(Comparison, ensembl_gene_id, logFC) %>%
  group_by(Comparison) %>%
  dplyr::summarise(FDR_0_05_FC_1.2_DOWN = length(unique(ensembl_gene_id[logFC < 0])),
                   FDR_0_05_FC_1.2_UP = length(unique(ensembl_gene_id[logFC > 0])))

kable(full_join(tmp,tmp1))

p = DE2 %>%
  ggplot(aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position = 'top')
p = p + facet_grid(Comparison~., scales = 'free_y')
p = p + ggtitle('Differential Expression')
p
```

### Gene set enrichment analysis
Using Fisher's exact test for gene set enrichment analysis
```{r enrich.de, fig.height=16, fig.width=16}
all.diff.exp = list(DE %>%
                      tidyr::unite(Comparison1, to.Dx, to.Tissue) %>%
                      tidyr::unite(Comparison2, ref.Dx, ref.Tissue) %>%
                      tidyr::unite(Comparison, Comparison1, Comparison2, sep = '-'), 
                    DE1%>%
                      tidyr::unite(Comparison1, to.Dx, to.Tissue, to.Gender) %>%
                      tidyr::unite(Comparison2, ref.Dx, ref.Tissue, ref.Gender) %>%
                      tidyr::unite(Comparison, Comparison1, Comparison2, sep = '-'), 
                    DE2) %>%
  data.table::rbindlist(fill = TRUE, use.names = TRUE)
diffexp.genes.up = all.diff.exp %>%
  dlply(.(Comparison), .fun = function(x){
    tmp = x$hgnc_symbol[x$adj.P.Val <= 0.05 & x$logFC > 0] %>% unique
    tmp = tmp[!is.na(tmp)]
  })
names(diffexp.genes.up) = paste(names(diffexp.genes.up), 'UP', sep = '.')

diffexp.genes.down = all.diff.exp %>%
  dlply(.(Comparison), .fun = function(x){
    tmp = x$hgnc_symbol[x$adj.P.Val <= 0.05 & x$logFC < 0] %>% unique
    tmp = tmp[!is.na(tmp)]
  })
names(diffexp.genes.down) = paste(names(diffexp.genes.down), 'DOWN', sep = '.')

diffexp.genes = c(diffexp.genes.up, diffexp.genes.down)
diffexp.genes = diffexp.genes[sapply(diffexp.genes, length) > 20]
backgroundGenes = unique(DE$hgnc_symbol)

# Download all related genesets from synapse
geneset.files = synQuery('select * from file where parentId == "syn3240583"')
ALL_USED_IDs = c(ALL_USED_IDs, geneset.files$file.id)
genesets.to.test = sapply(geneset.files$file.name, function(dataSource, geneset.files){
  tmp = downloadFile(geneset.files$file.id[geneset.files$file.name == dataSource]) %>%
    dlply(.(Name), .fun = function(x){
      str_split(x$symBeforeOverlap, '\\|') %>% unlist %>% unique
    })
  }, geneset.files) %>%
  CovariateAnalysis::filterGeneSets(backgroundGenes) %>%
  do.call(c,.)

# Perform enrichment analysis
enrichment.results = llply(diffexp.genes, .fun = function(x, genesetsToTest, backgroundGenes){
   ldply(genesetsToTest, .fun = function(y,x,backgroundGenes){
     CovariateAnalysis::fisherEnrichment(x, y, backgroundGenes)
    }, x, backgroundGenes, .parallel = TRUE) %>%
    dplyr::mutate(fdr = p.adjust(pval, method = 'fdr'))
}, genesets.to.test, backgroundGenes) %>%
  rbindlist(idcol = 'Comparison', use.names = T, fill = T) %>%
  dplyr::rename(Category = .id)
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  
  tissueTypeAbrv	= 'DLPFC, ACC', 
  study = 'MSSM', 

  organism = 'HomoSapiens',
  consortium	= 'CMC',
   
  normalizationStatus	= TRUE,
  
  normalizationType	= 'CQN'
)

# Code
CODE <- Folder(name = "ACC and DLPFC - STAR - FEATURE COUNTS - ENSEMBL v75 - CQN - MSSM", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'CMC_ACC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('CMC_ACC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
annotations(COV_OBJ) = annotations(CODE)
annotations(COV_OBJ)$dataSubType = 'covariates'
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('CMC_ACC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
CQN.GENE_EXPRESSION$y[is.na(CQN.GENE_EXPRESSION$E)] = NA
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('CMC_ACC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
CQN.GENE_EXPRESSION$offset[is.na(CQN.GENE_EXPRESSION$E)] = NA
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('CMC_ACC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store design matrix
DESIGN %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'CMC_ACC_DLPFC_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('CMC_ACC_DLPFC_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store residual gene expression for network analysis
RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_ACC_DLPFC_netResidualExpression.tsv', 
                name = 'Normalised, covariates removed residual expression (for network analysis)', 
                parentId = CODE$properties$id)
annotations(nEXP_OBJ) = annotations(CODE)
annotations(nEXP_OBJ)$dataSubType = 'residualGeneExpForNetAnlz'
nEXP_OBJ = synStore(nEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store differential expression results
write.table(all.diff.exp, file = 'CMC_ACC_DLPFC_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('CMC_ACC_DLPFC_DiffExpression.tsv', 
                name = 'Differential Expression Results (Dx.Tissue)', 
                parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store enrichment results
write.table(enrichment.results, file = 'CMC_ACC_DLPFC_EnrichResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('CMC_ACC_DLPFC_EnrichResults.tsv', 
                  name = 'Enrichment Analysis Results (Dx.Tissue)', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichResults'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Residual Expression (for network analysis) |  `r nEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Analysis                        |  `r ENRICH_OBJ$properties$id`    |

### Source Code
[R markdown](`r thisFile`)