---
title: "Covariate analysis of MAYO reprocessed transcripts (with TMM normalisation and covariates adjusted)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./MAYO_REPROCESSED_ISOFORM.Rmd",
                                 parentId = "syn8672415",
                                 entityName = "MAYO Reprocessed Isoform (Normalised)")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, cache=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(reshape2)
library(limma)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(RColorBrewer)

library(synapseClient)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

library(doParallel)
library(foreach)

cl = makeCluster(detectCores()-2)
doParallel::registerDoParallel(cl)

synapseLogin()

options(xtable.type="html")
source('../R/parallelDuplicateCorrelation.R')

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE}
# Synapse parameters
parentId = 'syn8672415'
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MAYO reprocessed transcript data';

thisFileName <- 'MAYO_REPROCESSED_ISOFORM.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD_isoform')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data}
# Get reprocessed sample counts for temporal cortex
COUNT_TC_ID <- 'syn10507725';
ALL_USED_IDs <- COUNT_TC_ID
COUNT_TC <- read.table(synGet(COUNT_TC_ID)@filePath, header=T, sep='\t', check.names = F, row.names = 1)

# Get reprocessed sample counts for cerebellum
COUNT_CE_ID <- 'syn10499229';
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_CE_ID
COUNT_CE <- read.table(synGet(COUNT_CE_ID)@filePath, header=T, sep='\t', check.names = F, row.names = 1)

COUNT = cbind(COUNT_TC, COUNT_CE[rownames(COUNT_TC),])

# Get clinical metadata
METADATA_TC_ID <- 'syn3817650'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TC_ID
METADATA_TC <- read.table(synGet(METADATA_TC_ID)@filePath,sep='\t',header=T, stringsAsFactors=F) %>%
  tidyr::separate(ID, c('Donor_ID', 'BrainRegion'), sep = '_') %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_'))

# Get clinical metadata
METADATA_CE_ID <- 'syn5223705'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_CE_ID
METADATA_CE <- read.table(synGet(METADATA_CE_ID)@filePath,sep='\t',header=T, stringsAsFactors=F) %>%
  tidyr::separate(SampleID, c('Donor_ID', 'BrainRegion')) %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_')) %>%
  dplyr::rename(Gender = Sex, FLOWCELL = Flowcell)

# Get PMI estimates from gene expression analysis
METADATA_PMI = downloadFile('syn8466814') %>%
  dplyr::select(SampleID, PMI) %>%
  dplyr::rename(ID = SampleID)
ALL_USED_IDs[length(ALL_USED_IDs)+1] = 'syn8466814'

# Merge all metadata
METADATA = rbindlist(list(METADATA_TC, METADATA_CE), use.names = T, fill = T) %>%
  left_join(METADATA_PMI) %>%
  as.data.frame

# Samples excluded by investigators
SAMPLES.EXCLUDE = c(downloadFile('syn6126119')$`Sample Name`,
                    downloadFile('syn6126114')$`Sample Name`)
ALL_USED_IDs = c(ALL_USED_IDs, 'syn6126119', 'syn6126114')
```
Following `r length(SAMPLES.EXCLUDE)` samples were removed from analysis as original investigators tagged them as outliers for various reasons as described here syn6126119, syn6126114
Samples removed are `r paste(SAMPLES.EXCLUDE, collapse = ', ')`

### Data preprocessing
```{r preprocess.data, echo=TRUE}
# Fix metadata
METADATA = METADATA %>%
  dplyr::mutate(AgeAtDeath = gsub("_or_above", "", AgeAtDeath),
                Diagnosis = gsub("Pathologic Aging", "OTHER", Diagnosis),
                Diagnosis = gsub("PSP", "OTHER", Diagnosis),
                Diagnosis = factor(Diagnosis, 
                                   levels = c("AD", "Control", "OTHER")),
                BrainRegion.Diagnosis = paste(BrainRegion, Diagnosis, sep = '.'),
                RIN2 = RIN^2,
                ApoE = factor(ApoE)) %>%
  filter(!is.na(Source)) %>%
  dplyr::filter(!(ID %in% SAMPLES.EXCLUDE))
levels(METADATA$ApoE) = c('0','0','1','0','1','2')
METADATA$BrainRegion.ApoE = paste(METADATA$BrainRegion, METADATA$ApoE, sep = '.')
```

```{r preprocess.data1}
# Match covariates to expression data
indToRetain = intersect(METADATA$ID, colnames(COUNT))
indRemoved = setdiff(colnames(COUNT), METADATA$ID)

COUNT <- COUNT[, indToRetain]

rownames(METADATA) = METADATA$ID
METADATA = METADATA[indToRetain,]
```
`r dim(COUNT)[2]` samples from `r length(unique(METADATA$Donor_ID))` subjects were obtained from the Mayo cohorts in AMP-AD reprocessed rnaseq project. 

Following samples were removed `r indRemoved`

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering}
primaryVariable <- c("BrainRegion.Diagnosis", "BrainRegion.ApoE")
FactorCovariates <- c("Donor_ID", "Source", "Gender", "FLOWCELL", "BrainRegion.Diagnosis", "BrainRegion.ApoE")
ContCovariates <- c("RIN", "RIN2", "AgeAtDeath","PMI")
  
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$ID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=6}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm'))
```

```{r data.explore, fig.width = 12, fig.height = 5}
my.theme = theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = BrainRegion.Diagnosis, y = RIN)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme

# AgeAtDeath
p[[2]] = ggplot(COVARIATES, aes(x = BrainRegion.Diagnosis, y = as.numeric(AgeAtDeath))) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeAtDeath') + my.theme

# PMI
p[[3]] = ggplot(COVARIATES, aes(x = BrainRegion.Diagnosis, y = as.numeric(PMI))) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + my.theme

multiplot(plotlist = p, cols = 3)
```
### Filter transcripts
Remove genes that have less than 5 cpm counts in at least 50% of samples per BrainRegion x Diagnosis
```{r cpmnormalisation}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('SampleID') %>%
  dlply(.(BrainRegion.Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                   MIN_GENE_CPM=5, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

### Outlier Analysis
```{r outlier.analysis, cache = FALSE, fig.width = 10}
indToRemove = c("11311_CER", "1923_CER", "1923_TCX", "11396_TCX", "11294_TCX", "11408_TCX",
                '1950_TCX', '1950_CER', '1925_TCX', '1957_CER')

# Find principal components of expression to plot
PC <- prcomp(voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(label = SampleID) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.')
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="top") 
p <- p + facet_grid(BrainRegion~.+Diagnosis, scales = 'free')
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.')

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(BrainRegion~.+Diagnosis)
p

indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]

tmp = COVARIATES %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  group_by(BrainRegion, Diagnosis) %>%
  summarise(count = n()) %>%
  spread(Diagnosis, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

Distribution of samples are: `r kable(tmp)`

### Winsorise counts (Gene outliers in samples)
```{r winsorise.data}
LOG.CPM = cpm(PROCESSED_COUNTS$filteredExprMatrix$counts, log = T)

# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(LOG.CPM, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  
  ind.low = which(x == max(x[which(x <= mn-3*std.dev)], na.rm = T))
  ind.high = which(x == min(x[which(x >= mn+3*std.dev)], na.rm = T))
  
  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t

# Back calculate counts for further analysis
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Library Normalisation
Initial library normalisation usign tmm
```{r tmm}
# Compute offset for gene length and gc content
TMM.GENE_EXPRESSION = calcNormFactors(NEW.COUNTS, method = 'TMM')
VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, design = NULL, lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)
  
# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(VOOM.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  
  ind.low = which(x == max(x[which(x <= mn-3*std.dev)], na.rm = T))
  ind.high = which(x == min(x[which(x >= mn+3*std.dev)], na.rm = T))
  
  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t
VOOM.GENE_EXPRESSION$E = LOG.CPM
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Sample clustering
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.')

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="top") + facet_grid(BrainRegion+.~Diagnosis, scales = 'free_y')
p
```
Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
COVARIATES.tmp = (COVARIATES[,c(FactorCovariates), drop = F])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(VOOM.GENE_EXPRESSION$E)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```
```{r temp, include = F}
dev.off()
gc()
```
### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(VOOM.GENE_EXPRESSION$E, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix.

NOTE:
1. Using a mixed effect model where random effect is chosen as Donor_ID
2. Adding Source, FLOWCELL and Gender a priori to variable selection
3. Primary variable of interest BrainRegion.Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Source', 'FLOWCELL', 'Gender');

# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, 
                         c(postAdjustCovars, primaryVariable, 'Donor_ID'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
notEstimable = c()

loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects and Donor_ID as random effect'))

  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

  # Estimate voom weights for dispersion control
  VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, design=DM1$design, plot=F, lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

  # Calculate correlation values of random effects
  VOOM.GENE_EXPRESSION$E = VOOM.GENE_EXPRESSION$E
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$Donor_ID,
                                             method = 'lmer')
  # correlation = list(); correlation$cor = 0.457197
  
  if (!is.nan(correlation$cor)){
    # Re-estimate voom weights
    VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, 
                                design=DM1$design, plot=F, 
                                block = COVARIATES$Donor_ID, 
                                correlation = correlation$cor)

    # Fit linear model using new weights and new design
    VOOM.ADJUSTED.FIT = lmFit(VOOM.GENE_EXPRESSION$E,
                              design = DM1$design,
                              weights = VOOM.GENE_EXPRESSION$weights,
                              block = COVARIATES$Donor_ID, 
                              correlation = correlation$cor)

    # Residuals after normalisation
    RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                  VOOM.GENE_EXPRESSION$E)
  } else {
    notEstimable = c(notEstimable, postAdjustCovars[length(postAdjustCovars)])
    postAdjustCovars = postAdjustCovars[1:(length(postAdjustCovars)-1)]
  }
  
  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates),
                         c(postAdjustCovars, primaryVariable, 'Donor_ID', notEstimable))

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  residCovars = setdiff(residualSigCovars$significantCovars, c(postAdjustCovars, primaryVariable, notEstimable))
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}

modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')

tmp <- paste('Using following covariates in the final model:', modelStr)
```

### Sanity check
```{r residual.adj, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)

residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2.1, fig.height=10, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('Region','Diagnosis'), sep = '\\.') %>% 
  dplyr::mutate(Region = factor(Region),
                Source = factor(Source))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Region~., scales = 'free_y')
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
p
```

```{r decompse.normalise.data2.2, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,FactorCovariates])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

```{r temp1, include=F}
dev.off()
gc()
```

### Differential expression analysis (with BrainRegion.Diagnosis as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of BrainRegion.Diagnosis
1. CER.AD: Alzheimers Disease samples from Cerebellum region
2. CER.CONTROL: Control samples from Cerebellum region
3. CER.OTHER: Progressive Supranuclear Palsy and Pathological Aging samples from Cerebellum region
4. TCX.AD: Alzheimers Disease samples from Temporal cortex region
5. TCX.CONTROL: Control samples from Temporal cortex region
6. TCX.OTHER: Progressive Supranuclear Palsy and Pathological Aging samples from Temporal cortex region

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp, fig.height=10, fig.width=18}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[1], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(NEW.COUNTS, design=DESIGN$design, plot=F, lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')  

# Re-estimate voom weights
VOOM.WEIGHTS = voom(NEW.COUNTS,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor,
                    lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)

# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.OTHER",
  "BrainRegion.DiagnosisCER.AD-BrainRegion.DiagnosisCER.OTHER",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.OTHER",
  "BrainRegion.DiagnosisTCX.AD-BrainRegion.DiagnosisTCX.OTHER"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Tx.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('BrainRegion.Diagnosis','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrainRegion.ref','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('BrainRegion.ag','against'), sep = '\\.') %>%
  tidyr::unite(Comparison, reference, against, sep = '-')
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(Tx.ID, Comparison, Direction, BrainRegion.ref) %>%
  group_by(BrainRegion.ref, Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(Tx.ID))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(BrainRegion.ref+.~Comparison, scales = 'fixed')
p

all.diff.exp = list(Diagnosis = DE)
all.fit = list(Diagnosis = FIT)
```

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=10, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE %>%
                   dplyr::filter(Comparison == 'Control-AD'), 
                 aes(x = log10(gene.length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(gene.length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(BrainRegion.ref+.~Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')

pl[[2]] = ggplot(DE %>%
                   dplyr::filter(Comparison == 'Control-AD'), 
                 aes(x = percentage_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(BrainRegion.ref+.~Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')

pl[[3]] = ggplot(DE %>%
                   dplyr::filter(Comparison == 'Control-AD'), 
                 aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(BrainRegion.ref+.~Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')

multiplot(plotlist = pl, cols = 3)
```

### Differential expression analysis (with BrainRegion.ApoE as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of BrainRegion.Diagnosis
1. CER.0: No ApoE4 allele in cerebellum
2. CER.1: 1 ApoE4 allele in cerebellum
3. CER.2: 2 ApoE4 allele in cerebellum
4. TCX.0: No ApoE4 allele in temporal cortex
5. TCX.1: 1 ApoE4 allele in temporal cortex
6. TCX.2: 2 ApoE4 allele in temporal cortex

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp1, fig.height=10, fig.width=18}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(NEW.COUNTS, design=DESIGN$design, plot=F, lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')  

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E

# Re-estimate voom weights
VOOM.WEIGHTS = voom(NEW.COUNTS,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor,
                    lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)

# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.ApoECER.2-BrainRegion.ApoECER.0",
  "BrainRegion.ApoECER.2-BrainRegion.ApoECER.1",
  "BrainRegion.ApoECER.1-BrainRegion.ApoECER.0",
  "BrainRegion.ApoETCX.2-BrainRegion.ApoETCX.0",
  "BrainRegion.ApoETCX.2-BrainRegion.ApoETCX.1",
  "BrainRegion.ApoETCX.1-BrainRegion.ApoETCX.0"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Tx.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('BrainRegion.ApoE','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrainRegion.ref','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('BrainRegion.ag','against'), sep = '\\.') %>%
  dplyr::mutate(reference = paste0('ApoE',reference),
                against = paste0('ApoE',against)) %>%
  tidyr::unite(Comparison, reference, against, sep = '-')
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(Tx.ID, Comparison, Direction, BrainRegion.ref) %>%
  group_by(BrainRegion.ref, Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(Tx.ID))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(BrainRegion.ref+.~Comparison, scales = 'fixed')
p

all.diff.exp = c(all.diff.exp, list(ApoE = DE))
all.fit = c(all.fit, list(ApoE = FIT))
```

### Differential expression analysis (with BrainRegion.Diagnosis x Gender as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp2, fig.height=5, fig.width=18}
# Get design matrix
COVARIATES$BrainRegion.Diagnosis.Gender = paste(COVARIATES$BrainRegion.Diagnosis, COVARIATES$Gender, sep = '.') %>%
  factor()
DESIGN = getDesignMatrix(COVARIATES[, c('BrainRegion.Diagnosis.Gender', setdiff(postAdjustCovars, 'Gender')), drop = F], 
                         Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(NEW.COUNTS, design=DESIGN$design, plot=F, lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')  

# Re-estimate voom weights
VOOM.WEIGHTS = voom(NEW.COUNTS,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor,
                    lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)

# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.Diagnosis.GenderCER.AD.F-BrainRegion.Diagnosis.GenderCER.Control.F",
  "BrainRegion.Diagnosis.GenderCER.AD.M-BrainRegion.Diagnosis.GenderCER.Control.M",
  
  "BrainRegion.Diagnosis.GenderTCX.AD.F-BrainRegion.Diagnosis.GenderTCX.Control.F",
  "BrainRegion.Diagnosis.GenderTCX.AD.M-BrainRegion.Diagnosis.GenderTCX.Control.M"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Tx.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('BrainRegion.Diagnosis.Gender','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrainRegion.ref','Dx.ref','Gender.ref'), sep = '\\.') %>%
  tidyr::separate(to.state, c('BrainRegion.ag','Dx.ag','Gender.ag'), sep = '\\.') %>%
  tidyr::unite(Comparison, Dx.ref, Dx.ag, sep = '-')
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(Tx.ID, Comparison, Direction, BrainRegion.ref, Gender.ref) %>%
  group_by(BrainRegion.ref, Gender.ref, Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(Tx.ID))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Comparison+.~BrainRegion.ref+Gender.ref, scales = 'fixed')
p

all.diff.exp = c(all.diff.exp, list(Diagnosis.Gender = DE))
all.fit = c(all.fit, list(Diagnosis.Gender = FIT))
```

### Differential expression analysis (with BrainRegion.Diagnosis x AgeAtDeath as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp3, fig.height=5, fig.width=18}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c('BrainRegion.Diagnosis', setdiff(postAdjustCovars, 'AgeAtDeath')), drop = F], 
                         Intercept = F)
DESIGN$design[,grep('BrainRegion.Diagnosis',colnames(DESIGN$design))] = DESIGN$design[,grep('BrainRegion.Diagnosis',colnames(DESIGN$design))] * COVARIATES[,'AgeAtDeath']
colnames(DESIGN$design)[grep('BrainRegion.Diagnosis',colnames(DESIGN$design))] = 
  paste(colnames(DESIGN$design)[grep('BrainRegion.Diagnosis',colnames(DESIGN$design))],'AOD', sep = '.')
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(NEW.COUNTS, design=DESIGN$design, plot=F, lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')  

# Re-estimate voom weights
VOOM.WEIGHTS = voom(NEW.COUNTS,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor,
                    lib.size = TMM.GENE_EXPRESSION * LIB.SIZE)

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = VOOM.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)

# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.DiagnosisCER.AD.AOD-BrainRegion.DiagnosisCER.Control.AOD",
  "BrainRegion.DiagnosisCER.AD.AOD-BrainRegion.DiagnosisCER.OTHER.AOD",
  "BrainRegion.DiagnosisCER.OTHER.AOD-BrainRegion.DiagnosisCER.Control.AOD",
  "BrainRegion.DiagnosisTCX.AD.AOD-BrainRegion.DiagnosisTCX.Control.AOD",
  "BrainRegion.DiagnosisTCX.AD.AOD-BrainRegion.DiagnosisTCX.OTHER.AOD",
  "BrainRegion.DiagnosisTCX.OTHER.AOD-BrainRegion.DiagnosisTCX.Control.AOD"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = T) %>%
    rownameToFirstColumn('Tx.ID')
}, FIT.CONTR)
names(DE) = colnames(contrast)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('BrainRegion.Diagnosis','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrainRegion.ref','Dx.ref','AOD.ref'), sep = '\\.') %>%
  tidyr::separate(to.state, c('BrainRegion.ag','Dx.ag','AOD.ag'), sep = '\\.') %>%
  tidyr::unite(Comparison, Dx.ref, Dx.ag, sep = '-')
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(Tx.ID, Comparison, Direction, BrainRegion.ref, AOD.ref) %>%
  group_by(BrainRegion.ref, AOD.ref, Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(Tx.ID))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Comparison+.~BrainRegion.ref+AOD.ref, scales = 'fixed')
p

all.diff.exp = c(all.diff.exp, list(Diagnosis.AOD = DE))
all.fit = c(all.fit, list(Diagnosis.AOD = FIT))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Function to change entity ACL (also creates new acls if it doesn't exist)
synChangeEntityACL <- function(id, new.racl){
  tryCatch({ 
    # If old acl exist update
    old.acl = synGetEntityACL(id)
    old.acl@resourceAccess = new.racl
    synUpdateEntityACL(old.acl)
  }, error = function(e){ 
    # If not create a new acl
    new.acl = AccessControlList(id = id, resourceAccess = new.racl)
    synCreateEntityACL(new.acl)
  })
}

# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'isoformExp',
  summaryLevel = 'transcript',
  assay	 = 'RNAseq',
  assaySubType = 'Reprocessed',
  
  tissueTypeAbrv	= 'CBE, TCX', 
  study = 'MayoRNASeq', 
  center = 'UFL-MAYO-ISB',
  
  organism = 'HomoSapiens',
  consortium	= 'AMP-AD',
   
  normalizationStatus	= TRUE,
  modelSystem	= FALSE,
  
  normalizationType	= 'TMM'
)

# Code
CODE <- Folder(name = "MAYO Reprocessed Isoform (Normalised)", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'MAYO_CBE_TCX_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('MAYO_CBE_TCX_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
annotations(COV_OBJ) = annotations(CODE)
annotations(COV_OBJ)$dataSubType = 'covariates'
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('Tx.ID') %>%
  write.table(file = 'MAYO_CBE_TCX_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MAYO_CBE_TCX_Counts.tsv', name = 'TPM (raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store filtered counts
NEW.COUNTS %>%
  rownameToFirstColumn('Tx.ID') %>%
  write.table(file = 'MAYO_CBE_TCX_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ECOUNT_OBJ = File('MAYO_CBE_TCX_eCounts.tsv', name = 'TPM (estimated)', parentId = CODE$properties$id)
annotations(ECOUNT_OBJ) = annotations(CODE)
annotations(ECOUNT_OBJ)$dataSubType = 'estimatedCounts'
ECOUNT_OBJ = synStore(ECOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
VOOM.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Tx.ID') %>%
  write.table(file = 'MAYO_CBE_TCX_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MAYO_CBE_TCX_logCPM.tsv', name = 'TPM (filtered)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store design matrix
DM1$design %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'MAYO_CBE_TCX_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('MAYO_CBE_TCX_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store differential expression results
rbindlist(all.diff.exp, use.names = T, fill = T, idcol = 'Model') %>%
  write.table(file = 'MAYO_CBE_TCX_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MAYO_CBE_TCX_DiffExpression.tsv', 
                name = 'Differential Expression Results (BrainRegion.Diagnosis)', 
                parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store all model fits
save(list = 'all.fit', file = 'MAYO_CBE_TCX_allfits.RData')
AFIT_OBJ = File('MAYO_CBE_TCX_allfits.RData', 
                name = 'All model fits (BrainRegion.Diagnosis)', 
                parentId = CODE$properties$id)
annotations(AFIT_OBJ) = annotations(CODE)
annotations(AFIT_OBJ)$dataSubType = 'modelFits'
AFIT_OBJ = synStore(AFIT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (estimated)                         |  `r ECOUNT_OBJ$properties$id`    |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
