---
title: "Covariate analysis of MAYO reprocessed counts (with TMM normalisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./Mayo_RNASeq_Reprocessed_TMM.Rmd",
                                 parentId = "syn5570325",
                                 entityName = 'UFL_Mayo_ISB Reprocessed RNASeq TMM')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, cache=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get it from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(reshape2)
library(limma)
library(Biobase)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(glmnet)

library(ComplexHeatmap)
library(circlize)

library(synapseClient)
library(knitr)
library(githubr)

library(parallel)
library(foreach)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

synapseLogin()

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE}
# Synapse parameters
parentId = 'syn5570325';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of UFL_Mayo_ISB RNASeq data';

thisFileName <- 'Mayo_RNASeq_Reprocessed_TMM.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Get reprocessed sample counts for temporal cortex
COUNT_TC_ID <- 'syn7987765';
ALL_USED_IDs <- COUNT_TC_ID
COUNT_TC <- read.table(synGet(COUNT_TC_ID)@filePath, header=T, sep='\t', check.names = F)
colnames(COUNT_TC) = paste(colnames(COUNT_TC), 'TCX', sep = '_')

# Get reprocessed sample counts for cerebellum
COUNT_CE_ID <- 'syn7513807';
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_CE_ID
COUNT_CE <- read.table(synGet(COUNT_CE_ID)@filePath, header=T, sep='\t', check.names = F)

COUNT = cbind(COUNT_TC, COUNT_CE[rownames(COUNT_TC),])

# Get clinical metadata
METADATA_TC_ID <- 'syn3817650'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TC_ID
METADATA_TC <- read.table(synGet(METADATA_TC_ID)@filePath,sep='\t',header=T, stringsAsFactors=F) %>%
  tidyr::separate(ID, c('Donor_ID', 'BrainRegion'), sep = '_') %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_'))

# Get clinical metadata
METADATA_CE_ID <- 'syn5223705'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_CE_ID
METADATA_CE <- read.table(synGet(METADATA_CE_ID)@filePath,sep='\t',header=T, stringsAsFactors=F) %>%
  tidyr::separate(SampleID, c('Donor_ID', 'BrainRegion')) %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_')) %>%
  dplyr::rename(Gender = Sex, FLOWCELL = Flowcell)

METADATA = rbindlist(list(METADATA_TC, METADATA_CE), use.names = T, fill = T) %>%
  as.data.frame
```

### Data preprocessing
```{r preprocess.data,cache=FALSE, echo=TRUE}
# Fix metadata
METADATA$AgeAtDeath = gsub("_or_above", "", METADATA$AgeAtDeath)
METADATA$Diagnosis = gsub("Pathologic Aging","Pathologic_Aging", METADATA$Diagnosis)
METADATA$BrainRegion.Diagnosis = paste(METADATA$BrainRegion, METADATA$Diagnosis, sep = '.')
METADATA$RIN2 = METADATA$RIN^2
METADATA <- METADATA %>%
  filter(!is.na(Source))
# %>%
#   dplyr::filter(!(ID %in% c("11311_CER", "1923_CER", "1923_TCX", 
#                             "11396_TCX", "11294_TCX", "11408_TCX")))
```
Following samples were considered outliers and removed based on PCA and/or eucledian distance: 11311\_CER, 1923\_CER, 1923\_TCX, 11396\_TCX, 11294\_TCX, 11408\_TCX
```{r preprocess.data1}
# Match covariates to expression data
indToRetain = intersect(METADATA$ID, colnames(COUNT))
indRemoved = setdiff(colnames(COUNT), METADATA$ID)

COUNT <- COUNT[, indToRetain]

rownames(METADATA) = METADATA$ID
METADATA = METADATA[indToRetain,]
```

`r dim(COUNT)[2]` samples from `r length(unique(METADATA$Donor_ID))` subjects were obtained from the Mayo cohorts in AMP-AD reprocessed rnaseq project. 

Following samples were removed `r indRemoved`

### Predict missing PMI
```{r predict.pmi}
PMI = METADATA[,c('PMI', 'ID'), drop = F]
rownames(PMI) = METADATA$ID

# Get genes that are predictors of PMI and use them as control probes
CONTROL.PROBES_ID = 'syn6145639'
ALL_USED_IDs = c(ALL_USED_IDs, CONTROL.PROBES_ID)
CONTROL.PROBES = fread(synGet(CONTROL.PROBES_ID)@filePath, data.table = F)

# Remove genes
EXPR <- getGeneFilteredGeneExprMatrix(COUNT, 
                                      MIN_GENE_CPM=0,
                                      MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0,
                                      verbose=T)

# TMM normalisation
TMM.GENE_EXPRESSION = calcNormFactors(EXPR$filteredExprMatrix, method = 'TMM')

# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=NULL, plot=F)

ind = rownames(VOOM.GENE_EXPRESSION$E) %in% CONTROL.PROBES$x 
EXPR = VOOM.GENE_EXPRESSION$E[ind,]
PMI = PMI[colnames(EXPR),,drop=F]

expr = dlply(PMI, .(ID), function(mtd, expr){
  if (dim(mtd)[1] > 1)
    expr = rowMeans(expr[, mtd$ID]) %>% rownameToFirstColumn('GeneID')
  else
    expr = expr[, mtd$ID] %>% rownameToFirstColumn('GeneID')
  expr = expr %>% plyr::rename(c('DF' = unique(mtd$ID)))
  return(expr)
}, EXPR) %>%
  join_all
rownames(expr) = expr$GeneID

PMI = PMI[,colnames(PMI) %in% c('ID','PMI')] %>% unique
rownames(PMI) = PMI$ID

PMI = PMI[intersect(rownames(PMI), colnames(expr)), ]
expr = expr[,intersect(rownames(PMI), colnames(expr))]

# Split data into trainining and prediction set
x.train = expr[,!is.na(PMI$PMI)]
x.predict = expr[,is.na(PMI$PMI)]

y.train = PMI[!is.na(PMI$PMI),]

plr.model = cv.glmnet(t(x.train), y.train$PMI, family = 'gaussian')
coeff = as.matrix(coef.glmnet(plr.model, s = 'lambda.min'))
mse = sum((predict(plr.model, t(x.train), s = 'lambda.min', type = 'response') - y.train$PMI)^2)/length(y.train$PMI)
y.predict = predict(plr.model, t(x.predict), s = 'lambda.min', type = 'response') %>%
  rownameToFirstColumn('ID') %>%
  plyr::rename(c('1' = 'PMI'))

PMI = rbindlist(list(y.train, y.predict), use.names = T, fill = T)

METADATA = METADATA %>% 
  dplyr::select(-PMI) %>%
  left_join(PMI)
```

```{r diagnosisPlots, fig.height=6, fig.width = 10}
# RIN
p = list()
p[[1]] = ggplot(METADATA, aes(x = BrainRegion.Diagnosis, y = RIN, fill = Tissue)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

# AgeAtDeath
p[[2]] = ggplot(METADATA, aes(x = BrainRegion.Diagnosis, y = as.numeric(AgeAtDeath), fill = Tissue)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeAtDeath') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

# PMI
p[[3]] = ggplot(METADATA, aes(x = BrainRegion.Diagnosis, y = as.numeric(PMI), fill = Tissue)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

multiplot(plotlist = p, cols = 3)

# Source
vcd::mosaic(~ Source + Diagnosis + Tissue, data = METADATA)
```

### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c("Donor_ID", "Source", "Gender", "FLOWCELL", "BrainRegion.Diagnosis")
ContCovariates <- c("RIN", "RIN2", "AgeAtDeath","PMI")
  
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$ID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Covariate correlation
```{r covariates.correlation, fig.width=8, fig.height=8}
COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left')
```
### Gene Filtering
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis. 

Remove genes that have less than 1 cpm counts in at least 50% of samples per BrainRegion.Diagnosis
```{r cpmnormalisation}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(BrainRegion.Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
    }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% unique()

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = summary(COVARIATES$BrainRegion.Diagnosis) %>%
  rownameToFirstColumn('SampleName') %>%
  tidyr::separate(SampleName, c('BrainRegion','Diagnosis'), sep = '\\.') %>%
  tidyr::spread(Diagnosis, DF)
kable(tmp)
```

### Normalisation (with NULL)
Initial normalisation usign voom (with NULL design)
```{r initial.voom.normalisation, fig.height=5, fig.width=5}
# TMM normalisation
TMM.GENE_EXPRESSION = calcNormFactors(PROCESSED_COUNTS$filteredExprMatrix, method = 'TMM')

# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=NULL, plot=T)

# Set gene counts in specific samples that are deviating 3 sd from other samples to NA
log.mat = apply(VOOM.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  return((x < (mn-3*std.dev)) | (x > (mn+3*std.dev)))
}) %>% t
PROCESSED_COUNTS$filteredExprMatrix[log.mat] = NA

# TMM normalisation
TMM.GENE_EXPRESSION = calcNormFactors(PROCESSED_COUNTS$filteredExprMatrix, method = 'TMM')

# Initial normalisation of gene expression
VOOM.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=NULL, plot=T)
```

Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data1, fig.height=14, fig.width=9, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale. = T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(ID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2],
                       PC3=PC$rotation[,3],
                       PC4=PC$rotation[,4])

plotdata <- left_join(plotdata, METADATA, by="ID")

pl = list()

plotdata[grep('CER', plotdata$ID),'Region'] = 'CER'
plotdata[grep('TCX', plotdata$ID),'Region'] = 'TCX'
plotdata$Region = factor(plotdata$Region)
plotdata$Source = factor(plotdata$Source)
plotdata$ID[!(plotdata$ID %in% c("11311_CER", "1923_CER", "1923_TCX", "11396_TCX", "11294_TCX", "11408_TCX"))] = ''

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape = Source, color = Region))
p <- p + theme_bw() + theme(legend.position = "right") +scale_shape_manual(values = 1:2)
p <- p + facet_grid(Region~.)
p <- p + geom_text(aes(label= ID), size=4, hjust=1, vjust=-1)
p
```
```{r decompse.normalise.data2, fig.height=9, fig.width=25, results='asis'}
# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(VOOM.GENE_EXPRESSION$E)), method = "average")
tree = WGCNA::cutreeStatic(sampleTrees, cutHeight = 135, minSize=10)

COVARIATES.tmp = COVARIATES
COVARIATES.tmp[,FactorCovariates] = lapply(COVARIATES.tmp[,FactorCovariates], as.numeric)
col.mat = labels2colors(cbind(tree,COVARIATES.tmp))
col.mat = gsub('\\.1', '', col.mat)

WGCNA::plotDendroAndColors(sampleTrees, 
                           col.mat,
                           groupLabels = c('Cluster',colnames(COVARIATES.tmp)))

tmp = paste('Samples in cluster 0 are', paste(sampleTrees$labels[tree == 0], collapse = ','))
tmp = base::gsub('_','\\\\_',tmp)
writeLines(tmp)
```

```{r temp, include=FALSE}
dev.off()
gc()
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, results='asis', echo=FALSE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(VOOM.GENE_EXPRESSION$E, COVARIATES,
                                                 'NULL design(voom-normalized)',
                                                 isKeyPlot=TRUE)

# Find significant covariates
adjustCovars = preAdjustedSigCovars$significantCovars
```
Significant covariates to adjust at FDR 0.1 are `r paste(gsub('_','\\\\_',adjustCovars), collapse= ',')`
```{r preAdjustedSigCovars.NULL.ALL, fig.width=25, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (iterative)
Since many covariates are correlated, re-normalising COVARIATES with an iterative design matrix.
```{r iterative.adjusted.voom.normalisation, results='asis'}
postAdjustCovars = c('BrainRegion.Diagnosis', 'Source', 'FLOWCELL', 'Gender')
randomCovariates = 'Donor_ID'

# Assign residual covariates
residualSigCovars = preAdjustedSigCovars
significantCovars = residualSigCovars$significantCovars
covariatesEffects = residualSigCovars$Effects.significantCovars
covariatesEffects = covariatesEffects[setdiff(significantCovars, c(postAdjustCovars, randomCovariates))]

postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))

loopCount = 0 
while(length(significantCovars)!=0 && loopCount <= 100){
  writeLines(paste('Using following covariates in the model:', 
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))
  
  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]
    
  # Estimate voom weights
  VOOM.ADJUSTED.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=DM1$design, plot=F)
  
  # Calculate correlation values of random effects
  correlation = parallelDuplicateCorrelation(VOOM.ADJUSTED.GENE_EXPRESSION, block = COVARIATES$Donor_ID)
  
  # Re-estimate voom weights
  VOOM.ADJUSTED.GENE_EXPRESSION = voom(TMM.GENE_EXPRESSION, design=DM1$design, plot=F, 
                                       block = COVARIATES$Donor_ID, correlation = correlation$cor)
  
  # Fit a mixed effect linear model
  FIT = lmFit(VOOM.ADJUSTED.GENE_EXPRESSION)
  
  # Residuals after normalisation
  RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(FIT,VOOM.ADJUSTED.GENE_EXPRESSION$E)
  
  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates), c(postAdjustCovars, randomCovariates))
  
  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                                COVARIATES,
                                                'all adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)
  
  # Add postadjusted covariates (if any)
  significantCovars = residualSigCovars$significantCovars
  covariatesEffects = residualSigCovars$Effects.significantCovars
  covariatesEffects = covariatesEffects[setdiff(significantCovars,c(postAdjustCovars, randomCovariates))]
  
  postAdjustCovars = c(postAdjustCovars,names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}
modelStr <-paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=','), 'as fixed effects')

tmp <- paste('Using following covariates in the final model:', modelStr, 'and Donor_ID as random effects')
```
`r tmp`
### Sanity Check
Residualise significant covariates using a linear model and find correlation between PCA of residuals with covariates
```{r residualSigCovars.ALL, fig.width=25, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
Adding "BrainRegion.Diagnosis" back to the residuals
```{r addVarsBack}
varsToAddBackIn <-  grep("BrainRegion.Diagnosis", colnames(FIT$coefficients), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  t(data.matrix(DM1$design[,varsToAddBackIn]) %*% t(FIT$coefficients[,varsToAddBackIn]))
```

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=10, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(RESIDUAL.GENE_EXPRESSION)), method = "average")

# Cluster expression
clust = cutreeStatic(sampleTrees, cutHeight = 100, minSize = 10)

# Plot first 2 PCs
plotdata <- data.frame(ID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2],
                       PC3=PC$rotation[,3], 
                       PC4=PC$rotation[,4],
                       clust = clust)

plotdata <- left_join(plotdata, METADATA, by="ID")

pl = list()
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=factor(Source), color=as.factor(Tissue), size=as.numeric(AgeAtDeath)))
p <- p + theme_bw() + theme(legend.position="right") +scale_shape_manual(values = 1:2)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
pl[[1]] <- p

p <- ggplot(plotdata, aes(x=PC3, y=PC4))
p <- p + geom_point(aes(shape=factor(Source), color=as.factor(Diagnosis), size=as.numeric(AgeAtDeath)))
p <- p + theme_bw() + theme(legend.position="right") +scale_shape_manual(values = 1:2)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
pl[[2]] <- p

multiplot(plotlist = pl, cols = 1)

names(clust) = sampleTrees$labels
clust = clust[rownames(COVARIATES)]
```
```{r decompse.normalise.data4, fig.height=10, fig.width=25, results='asis'}
# Tree based clustering
COVARIATES.tmp = COVARIATES
COVARIATES.tmp[, FactorCovariates] = sapply(COVARIATES.tmp[,FactorCovariates], as.numeric)
col.mat = labels2colors(cbind(tree,COVARIATES.tmp))
col.mat = gsub('\\.1', '', col.mat)

plotDendroAndColors(sampleTrees, 
                    colors = col.mat, 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES))
```

```{r temp1, include=FALSE}
gc()
dev.off()
```

```{r get.gene.symbols}
# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host = "dec2016.archive.ensembl.org")

# Get background gene ids
backgroundGenes = data.frame(gene.id = rownames(COUNT), ID = rownames(COUNT)) %>%
  tidyr::separate(ID, c('ensembl_gene_id','position'), sep = '\\.')

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
                       mart = mart)

# Add gene names
RESIDUAL.GENE_EXPRESSION <- rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'gene.id') %>%
  left_join(backgroundGenes) %>% 
  left_join(Ensemble2HGNC) %>%
  dplyr::select(gene.id, ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))

VOOM.ADJUSTED.GENE_EXPRESSION$E <- rownameToFirstColumn(VOOM.ADJUSTED.GENE_EXPRESSION$E, 'gene.id') %>%
  left_join(backgroundGenes) %>% 
  left_join(Ensemble2HGNC) %>%
  dplyr::select(gene.id, ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=20}
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.Pathologic_Aging",
  "BrainRegion.DiagnosisCER.Control-BrainRegion.DiagnosisCER.PSP",
  "BrainRegion.DiagnosisCER.Pathologic_Aging-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisCER.Pathologic_Aging-BrainRegion.DiagnosisCER.PSP",
  "BrainRegion.DiagnosisCER.PSP-BrainRegion.DiagnosisCER.AD",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.Pathologic_Aging",
  "BrainRegion.DiagnosisTCX.Control-BrainRegion.DiagnosisTCX.PSP",
  "BrainRegion.DiagnosisTCX.Pathologic_Aging-BrainRegion.DiagnosisTCX.AD",
  "BrainRegion.DiagnosisTCX.Pathologic_Aging-BrainRegion.DiagnosisTCX.PSP",
  "BrainRegion.DiagnosisTCX.PSP-BrainRegion.DiagnosisTCX.AD"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

# Get differnetial expression
DE = lapply(1:12, function(i, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION){
  topTable(FIT.CONTR, coef=i, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1], confint = T)
}, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION)
names(DE) = gsub('BrainRegion.Diagnosis','',colnames(contrast))

DE1 = plyr::ldply(DE, .id = 'Comparison') %>%
  dplyr::mutate(Comparison1 = Comparison) %>%
  tidyr::separate(Comparison1, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrainRegion','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('BrainRegion','against'), sep = '\\.') %>%
  tidyr::unite(Name, reference, against, sep = '-') %>%
  dplyr::mutate(Comparison = Name, Study = 'MAYO') %>%
  dplyr::rename(gene.id = genes, Region = BrainRegion) %>%
  dplyr::select(-Name) %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC)

p <- ggplot(DE1, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p <- p + geom_hline(yintercept = -log10(0.05), color = 'red')
p <- p + geom_vline(xintercept = log2(1.2), color = 'red')
p <- p + geom_vline(xintercept = -log2(1.2), color = 'red')
p <- p + facet_grid(Region~.+Comparison)
p

tmp = DE1 %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(Region, Comparison) %>%
  dplyr::summarize(nDiffExpGenes = length(unique(gene.id))) %>%
  spread(Region, nDiffExpGenes)
kable(tmp)
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Reprocessed RNASeq TMM', parentId = parentId)
annotations(CODE) = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'Reprocessed RNAseq',
  
  tissueTypeAbrv	= 'CBE, TCX', 
  study = 'MayoRNASeq', 
  center = 'UFL-MAYO-ISB',
  
  organism = 'HomoSapiens',
  consortium	= 'AMP-AD',
   
  normalizationStatus	= TRUE,
  modelSystem	= FALSE,
  
  normalizationType	= 'TMM'
)
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'Sampleid')
write.table(COVARIATES, file = 'MAYO_CBE_TCX_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('MAYO_CBE_TCX_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix %>%
  as.matrix() %>% rownameToFirstColumn('gene.id') %>%
  write.table(file = 'MAYO_CBE_TCX_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MAYO_CBE_TCX_Counts.tsv', name = 'Raw counts (filtered)', parentId = CODE$properties$id)
COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                     executed = thisFile, activityDescription = activityDescription)

# Store logCPM
write.table(VOOM.ADJUSTED.GENE_EXPRESSION$E, file = 'MAYO_CBE_TCX_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MAYO_CBE_TCX_logCPM.tsv', name = 'Raw counts (logCPM)', parentId = CODE$properties$id)
LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)

# Store design matrix
write.table(DM1$design, file = 'MAYO_CBE_TCX_Design.tsv', sep = '\t', row.names=T, quote=F)
DM_OBJ = File('MAYO_CBE_TCX_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
DM_OBJ = synStore(DM_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                  executed = thisFile, activityDescription = activityDescription)

# Store residual gene expression
write.table(RESIDUAL.GENE_EXPRESSION, file = 'MAYO_CBE_TCX_Expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('MAYO_CBE_TCX_Expression.tsv', name = 'Normalised Adjusted Residual Expression (Dx added)', 
               parentId = CODE$properties$id)
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store differential expression results
write.table(DE1, file = 'MAYO_CBE_TCX_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MAYO_CBE_TCX_DiffExpression.tsv', name = 'Differential Expression Analysis', parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)

stopCluster(cl)
```
|  *Results*                |  *SynapseID*                    |
|  -------                  |   ---------                     |
|  Covariates               |  `r COV_OBJ$properties$id`      |
|  Raw counts (logCPM)      |  `r LCOUNT_OBJ$properties$id`   |
|  Design Matrix            |  `r DM_OBJ$properties$id`       |
|  Residual Expression      |  `r EXP_OBJ$properties$id`      |
|  Differential Expression  |  `r DEXP_OBJ$properties$id`     |

### Source R Code
[This File](`r thisFile`)