---
title: "Covariate analysis of MSBB microarray data"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

synapseLogin()

knit2synapse::knitToFolderEntity(file = "./MSSM_Array.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB Microarray')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE}
## It is assumed your working directory is where this file

### Clear R console screen output
cat("\014") 

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

## Needs the dev branch
library(rGithubClient)

synapseLogin()

# source utility files from ../R/lib folder
file.sources = list.files('../R/lib',pattern="*.R", full.names=T)
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE}
# Synapse parameters
parentId = 'syn5570248';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of MSBB microarrays';

thisFileName <- 'MSSM_Array.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
Following brain regions were analysed for differential expression between case and control

| Brain Region                        | Synapse ID |
| ----------------------------------- | ---------- |
| middle temporal                     | syn3191101 |
| superior temporal                   | syn3191103 |
| inferior temporal                   | syn3191099 |
| precuneus superior parietal lobule  | syn3191119 |
| caudal anterior cingulate           | syn3191107 |
| frontal pole                        | syn3191095 |
| para hippocampal gyrus              | syn3191109 |

```{r fxns}
# Function to do covariates adjustments
covariatesAdjustments <- function(exprId, regionName, covId){
  writeLines(paste0('Working on ',regionName))
  
  # Download expression data
  expr <- read.table(synGet(exprId)@filePath, header=T, sep='\t', check.names = F)
  expr <- expr[expr$Gene.Symbol!='',]
  rownames(expr) <- expr$ID
  ANNOTATIONS <- dplyr::select(expr,ID,GB_ACC,Gene.Symbol,ENTREZ_GENE_ID)
  expr <- dplyr::select(expr,-c(ID,GB_ACC,Gene.Symbol,ENTREZ_GENE_ID))
  rownames(expr) = ANNOTATIONS$ID

  # Collapse probe ids to hgnc symbols
  Collapsed.obj <- collapseRows(expr,ANNOTATIONS$Gene.Symbol,ANNOTATIONS$ID)
  expr <- Collapsed.obj$datETcollapsed

  # Download covariates data
  covar <- fread(synGet(covId)@filePath,sep='\t',header=T,data.table=F)
  BrainBank <- covar$BrainBank
  covar <- dplyr::select(covar,-(BrainBank))
  covar <- dplyr::mutate(covar,Dx = ifelse(CDR<1,'Control','AD'))
  rownames(covar) <- BrainBank

  # Get covariates classification
  FactorCovariates <- c('Sex','Race','CDR','Braak','NP1','Dx')
  ContCovariates <- c('PMI','pH','Age','PLQ_Mn','NPrSum','NTrSum')

  # Match expression to covariates data
  ind <- match(colnames(expr),rownames(covar))
  expr <- expr[,!is.na(ind)]
  covar <- covar[ind[!is.na(ind)],]

  # Convert covariates to factors and numeric variables accordingly
  covar$Age = gsub('\\+','',covar$Age)
  covar[,FactorCovariates] <- lapply(covar[,FactorCovariates], as.factor)
  covar[,ContCovariates] <- lapply(covar[,ContCovariates], as.numeric)  

  # Find PC of gene expression and significant covariates that are highly correlated with PCs
  covar.tmp = covar
  covar.tmp[,FactorCovariates] <- lapply(covar.tmp[,FactorCovariates], as.numeric)
  preAdjustedSigCovars <- runPCAandPlotCorrelations(expr, covar.tmp, 'expression', isKeyPlot=TRUE)

  # Plot significant data
  print(preAdjustedSigCovars$PC_res[[2]]$plotData)
  writeLines(paste('Significant covariates are:',paste(preAdjustedSigCovars$significantCovars,collapse=',')))

  # Get design matrix
  DESIGN <- getDesignMatrix(covar[,'Dx',drop=F], Intercept = F)
  writeLines(paste('Fitting linear model with the following coefficients:',paste(colnames(DESIGN$design),collapse=',')))

  # Fit linear model
  fit <- lmFit(expr, design = DESIGN$design)

  # Calculate residuals matrix
  residualFit <- residuals.MArrayLM(fit, expr)
  residualFit = residualFit + fit$coefficients[,c('DxAD','DxControl')] %*% t(DESIGN$design[,c('DxAD','DxControl')])

  # Find correlation between residuals and covariates 
  postAdjustedSigCovars <- runPCAandPlotCorrelations(residualFit, covar.tmp,'residual expression', isKeyPlot=TRUE)

  if (length(postAdjustedSigCovars$significantCovars) != 0)
    writeLines(paste('Following coefficients have to be included in the model:',paste(postAdjustedSigCovars$significantCovars,collapse=',')))

  return(list(EXPR = expr,
              COVAR = covar,
              FIT = fit, 
              RESIDUAL = residualFit, 
              DESIGN = DESIGN, 
              ANNOTATIONS = ANNOTATIONS))
}
```
### Extract expression data and adjust covariates 
```{r covariate.analysis, cache=TRUE, fig.height=10, fig.width=15, echo=FALSE, results='asis'}
# SynIDs of expression and covariates matrix
EXP_ID <- c('Middle Temporal' = 'syn3191101',
            'Superior Temporal' = 'syn3191103',
            'Inferior Temporal' = 'syn3191099',
            'Superior Parietal Lobule' = 'syn3191119',
            'Caudal Anterior Cingulate' = 'syn3191107',
            'Parahippocampal Gyrus' = 'syn3191109',
            'Frontal Pole' = 'syn3191107')
COV_ID <- 'syn3205399'  

# Obtain adjusted expession
adjustedExpression <- mapply(covariatesAdjustments, EXP_ID, names(EXP_ID), MoreArgs= list(COV_ID), SIMPLIFY=F)
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE}
# Code
CODE <- Folder(name = "MSBB Microarray", parentId ="syn5570248")
CODE <- synStore(CODE)

ALL.OBJ = mapply(function(exp, regionName, parentId, used, executed, activityName, activityDescription){
  # Expression
  expr = rownameToFirstColumn(exp$EXPR, 'hgnc_symbols')
  write.table(expr, file='expression.tsv', sep = '\t', row.names=F, quote=F)
  obj = File('expression.tsv', 
             name = paste('Expression',regionName), 
             parentId = parentId)
  obj = synStore(obj,
             used = used, 
             activityName = activityName, 
             executed = executed, 
             activityDescription = activityDescription)
  
  # Covariates
  expr = rownameToFirstColumn(exp$COVAR, 'SampleId')
  write.table(expr, file='covariates.tsv', sep = '\t', row.names=F, quote=F)
  obj = File('covariates.tsv', 
             name = paste('Covariates',regionName), 
             parentId = parentId)
  obj = synStore(obj,
             used = used, 
             activityName = activityName, 
             executed = executed, 
             activityDescription = activityDescription)
},adjustedExpression, names(adjustedExpression), 
MoreArgs = list(CODE$properties$id, ALL_USED_IDs, thisFile, activityName, activityDescription), 
SIMPLIFY = F)       
```